### 2011-05-18 paper/paper.tex
Final paper

\documentclass[11pt]{article}
\usepackage{acl-hlt2011}
\usepackage{color}
\usepackage{graphicx}
\usepackage{times}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{algorithmic}
\usepackage{multirow}
\usepackage{url}
\DeclareMathOperator*{\argmax}{arg\,max}
\setlength\titlebox{6.5cm}    % Expanding the titlebox
\frenchspacing

\title{Towards an Automated Sportscasting System for Chess}
\author{
  Charles Chen \\
	UC Berkeley \\
	Berkeley, CA 94720 \\
  {\tt \small{charleschen@berkeley.edu}} \\
\And
  Richard Shin \\
	UC Berkeley \\
	Berkeley, CA 94720 \\
  {\tt \small{ricshin@berkeley.edu}} \\
\And
  Jinghao Yan \\
	UC Berkeley \\
	Berkeley, CA 94720 \\
  {\tt \small{jinghao@berkeley.edu}} \\
}

\begin{document}
\maketitle

\begin{abstract}
We investigate applications of statistical learning to automated 
sportscasting for chess. In particular, this project studies two methods 
to automatically identify high level structures in chess games using 
support vector machines. Our first classifier is based on manual feature 
selection, while our second classifier employs sparse auto-encoding to 
automatically generate a set of feature vectors for a chess game. We 
then compare the results of the two different approaches and showed 
improvements using auto-encoding techniques. Finally, we conclude with 
results, analysis of errors, and ideas for potential future work.
\end{abstract}

\section{Introduction}
Chess has a long history of intriguing enthusiasts due to its 
complexity. Even after years of play, each game is new challenge, a new 
problem to solve. Historically, chess has been prominent as a sport 
where the battles of grandmasters engulf thousands; given the highly 
intricate nature of chess (and similar abstract strategy games), experts 
have scrutinized these games to devise standard high-level structures 
like openings, strategies, tactics, and time advantage, which they use 
to explain maneuvers and blunders that occur in the games. Devotees of 
chess also attempt to improve their own performance in terms of these 
high-level structures, reading chess theorists' comments about games by 
high-ranking players.

In this paper, with the end goal of automatically producing textual 
commentary similar to those produced by experts for grandmaster games, 
we investigate whether we might automatically identify such high-level 
patterns in real games. Specifically, we want to see how statistical 
machine learning techniques can help us automatically identify such 
structures using a large number of real chess games as training data, 
where some of the games may be already annotated with these high-level 
structures, while others may not be.

While others have created highly rigid and rule-based approaches to 
annotating chess games~\cite{cambridge-chess-annotation}, we 
specifically sought to avoid such approaches which require a lot of 
manual engineering and chess-specific expertise in order to obtain good 
results. In particular, for chess, there exists a large number of 
computer chess tools like chess engines, which search the game tree to 
examine the breadth of options available and identify the outcomes of 
each. We would like our approaches to have general applicability to 
other domains, so we chose to specifically avoid using such tools as 
much as possible.

To that end, we investigate the use of unsupervised feature learning and 
deep learning techniques on transcripts of chess games to see if they 
could automatically identify the high-level structures that make up a 
chess game. Deep learning techniques, like the sparse stacked 
auto-encoder, learns a more efficient way to represent some input given 
a large set of examples. In other fields such as computer vision, 
researchers have found that these techniques can give state-of-the-art 
results rivaling or exceeding features hand-engineered by experts. While 
learning a more compact way to represent parts of chess games, 
auto-encoders may be able to discover high-level structures similar to 
those used by human experts.

As a comparison, we also implement a small number of chess-specific, 
hand-engineered features based on information from a chess engine, such 
as the number of moves available from a specific board position. Given 
feature vectors made from hand-labeled games, we train SVM classifiers 
to predict the label of an unlabeled feature vector. We compare its 
performance to automatically learned features to see which one can more 
successfully predict high-level structures.

To conduct our experiments, we collect several thousand ``tactical 
training positions'' organized by theme, intended for use by human chess 
players to learn how to recognize such situations and improve their 
playing skills. We also collect a large number of unannotated chess 
games (about 2.5 million) from various sources on the web for use with 
semi-supervised and unsupervised learning techniques. After training our 
models, for each of the positions in the training data labeled with a 
theme (or high-level structure), we attempt to predict the correct 
theme(s) for a position using the models.

Unfortunately, we found that our approaches did not deliver accurate 
results in our experiments. In particular, the auto-encoder's output did 
not help significantly as a useful criterion for distinguishing between 
different types of themes. We examine the types of correct predictions 
made by our classifiers and discuss how we may improve our results in 
the future.

\section{Related Work}

\paragraph{Deep learning.} These techniques attempt to capture complex 
relationships between input variables in order to transform the input 
into a different representation which better expresses the relevant 
structures that make up the input. Specifically, they compose many 
layers of non-linearities in order to compactly represent complex 
functions which can model these relationships. They have been used 
successfully for many different types of tasks, especially in computer 
vision~\cite{Vincent,Lee,Socher}.

\paragraph{Automated sportscasting.} Chen and Mooney~\shortcite{chen} 
present a commentator system that learns language from sportscasts of 
simulated RoboCup games. Their system, which is trained on ambiguous 
supervision in the form of textual human commentaries and simulation 
states of the soccer games, simultaneously tries to establish 
correspondences between the commentaries and the simulation states. 
Their work focuses on generating meaningful English commentary from 
relatively simple topics such as ``Player 1 kicked the ball to Player 
2'', while our work focuses less on text generation and more on 
identifying complex topics.

\paragraph{Automatic chess commentary.} There exist techniques for 
automatic annotation of chess games which are heavily dependent on 
rule-based logic and use of a chess-engine to determine which parts of a 
specific chess game are salient~\cite{cambridge-chess-annotation}. As a 
departure from the rule-based approach to automatic chess commentary, 
Guid et al.~\shortcite{Guid} discuss the use of argument based machine 
learning, which analyzes reasoning where arguments for and against a 
certain claim are produced and evaluated (in the context of chess, why a 
certain move was played over others.) Their system, much like the one 
presented in this paper, identifies complex positional features of a 
given chess game.

However, we identified some shortcomings of the AMBL system. First, it 
relies on experts to provide arguments as to why a certain position 
exemplifies (or doesn.t) a particular positional feature. While this is 
an improvement from defining a classification rule manually, it is 
infeasible to repeat this procedure multiple times. Second, the 
arguments that the experts are allowed to provide rely on a preset list 
of features.

\section{Data Sources and Processing}
We obtained both annotated and unannotated records of chess games from 
various web sites in the PGN format \cite{pgn}, a widely-used format for 
compactly representing chess games. A large number of games as well as 
software and tools that use the PGN format are freely available.

\begin{table*}[p]
\centering
\vspace{-0.7cm}
\begin{tabular}{lp{0.6\textwidth}}
\hline
\textbf{Name} & \textbf{Description} \\ \hline
\tt{attackPossible} & The current player has opportunity to attack. \\
\tt{backrankWeakness} & There exists a future checkmate by a rook or 
queen on the edges of the board. \\
\tt{bishopPair} & There exists a decisive strategic advantage for the 
player with two bishops. \\
\tt{clearance} & There exists a maneuver to break the pawn structure. \\ 
\tt{closure} & There exists a maneuver to fix the pawn structure. \\
\tt{decoy} & There exists a maneuver to distract a crucial piece. \\
\tt{defensePressure} & There exists a maneuver to break the opposing 
King.s defense. \\
\tt{deflection} & There exists a maneuver to deflect an attack.  \\ 
\tt{discoveredAttack} & There exists a maneuver to allow for a 
discovered attack (an attack that was blocked by a friendly piece.) \\
\tt{doubleAttack} & There exists a move that simultaneously attacks at 
least two crucial squares. \\
\tt{drawCombinationsStalemate} & There exists a combination that allows 
for decisive material advantage or immediate checkmate. \\
\tt{enclosedKings} & There exists a maneuver to trap the opposing King. 
\\
\tt{freeingBlockingSquares} & There exists a move to free a crucial 
square. \\
\tt{goodEndgameChances} & There exists a move to allow for a favorable 
endgame. \\
\tt{inBetweenMove} & There exists an interrupting move that allows for 
decisive material advantage of immediate checkmate. \\ 
\tt{interference} & There exists a maneuver to block a crucial piece. \\ 
\tt{interruption} & There exists a maneuver that interrupts the 
opponent.s tactics. \\
\tt{kingInCenter} & There exists a decisive strategic advantage for the 
player with the King in the center of the board. \\
\tt{kingToBadSquares} & There exists a combination that forces the 
opposing King to move to a bad square. \\
\tt{longDiagonals} & There exists a strategic advantage for the player 
that controls squares on the long diagonal. \\
\tt{matingAttack} & There exists a decisive mating attack. \\
\tt{openingClosingDiagonals} & There exists a maneuver to clear a 
diagonal. \\
\tt{openingClosingFiles} & There exists a maneuver to clear a file. \\
\tt{overload} & There exists a combination that makes a defending piece 
defend too many squares at once. \\
\tt{passedPawn} & There exists a combination that allows for a passed 
pawn (a pawn with no opposing pawns in front of it.) \\
\tt{pawnPromotion} & There exists a combination that allows for a pawn 
promotion. \\
\tt{perpetualCheck} & There exists a combination that forces a perpetual 
check/draw. \\
\tt{pin} & There exists a combination that allows for a pin. \\
\tt{pullingPiecesToBadSquares} & There exists a combination that forces 
an opposing piece to move to a bad square. \\
\tt{queenCapture} & There exists a combination to capture the opposing 
Queen. \\
\tt{quietMates} & There exists a quiet move (one that looks 
inconsequential on the surface) leading to mate. \\
\tt{removalOfDefence} & There exists a maneuver to remove an opponent.s 
defender. \\
\tt{unpinning} & There exists a maneuver to unpin a piece. \\
\tt{wrongBishop} & The opponent.s only bishop is limited in use because 
of the pawn structure of the game. \\ \hline
\end{tabular}
\caption{List of high-level themes we attempt to automatically identify, 
along with explanations}
\label{table:themes}
\end{table*}

The already-annotated games come from databases of short snippets of 
chess 
games,\footnote{\url{http://wwwu.uni-klu.ac.at/gossimit/c/tactic.htm}} 
\footnote{\url{http://webplaza.pt.lu/public/ckaber/Chess.htm#Chess 
training}} intended for chess players to learn from by example, which 
are organized by particular themes that the students are to identify. 
These provided a useful repository of both interesting high-level 
concepts we could identify (we provide a listing in 
Table~\ref{table:themes}), as well as boards and moves corresponding to 
each one.

We also looked for as many chess games as possible from a variety of web 
sites, such as \url{chessgames.com}. We easily obtained over five 
million untagged games spread over 600,000 files and 5 GB of data. The 
number of annotated game snippets we obtained was slightly over 5,000.

\subsection{Data Pre-Processing}	
In order to ensure consistency and integrity over the games obtained 
from disparate sources, we sanitized them in multiple ways. We parsed 
each game in the data set and removed information that would confuse the 
parser. For example, some games had English or German comments by humans 
which we did not plan to use, and sometimes resulted in parsing errors 
due to specific characters present in the comments. Others had alternate 
branches, sometimes up to multiple levels (alternate branches within 
alternate branches), which the parser could not handle properly. Still 
others represented games in a non-standard way (for example, pawn 
promotions were represented without the ``=''), which we had to fix.

Since we obtained the games from many sources, there was significant 
overlap in games (approximately 50\% of the unlabeled games collected 
were redundant). Therefore, we also went through all of the downloaded 
games and eliminated duplicate games: multiple games that had the same 
initial position and move sequence were considered redundant and only 
one of each were kept.

\subsection{Parsing}
We tried various open source PGN parsers, each succumbing to flaws like 
mishandling uncommon cases (such as  promotions and special moves like 
\emph{en passant}). We based our system on 
Chesspresso,\footnote{\url{http://www.chesspresso.org/}} a Java library 
for writing chess programs, which we fixed and modified for our needs. 
Chesspresso takes a PGN file as input and returns a data structure that 
represents the initial position of the board and the subsequent moves 
(as well as all metadata necessary to recreate the PGN file). We also 
extended it to be more robust to variations in input format, in addition 
to being able to export PGNs in a canonical form, which was crucial for 
removing duplicates.

\subsection{Preparing the Data}
After parsing each game, we filtered out the tagged games that had more 
than 10 plies because it will be very difficult to identify the tags 
that apply to a complex sequence of moves. Among the games that we kept, 
we also kept only the first few positions and moves (the ``prefix'' of 
each game, as described earlier), a parameter that can be adjusted; for 
our experiments, we kept just the initial board specified in each game. 
We then generate labels for each game based on the information from the 
source websites. For example, games in a file that corresponds to a 
lesson called ``Double Attack'' are labeled {\tt doubleAttack}. In 
addition, some of those games may appear in other lessons, and 
consequently will have multiple labels. We also segment the games 
further by the color of the player that makes the first move in the 
game, in addition to the labels of the game, since most of the labels in 
Table~\ref{table:themes} only refer to a specific player; for instance, 
some of the games labeled with {\tt backrankWeakness} may have the white 
player with the weakness, while others show the black player with the 
weakness.

\section{Feature Extraction}
\subsection{Sparse auto-encoder}
The sparse auto-encoder is one of many methods in use for unsupervised 
feature learning~\cite{Coates}. It takes the form of a feedforward 
neural network, which consists of an input layer of N nodes, some number 
of hidden layers, and output layer of M nodes. The value of each layer 
$\mathbf{l_i}$ is determined with a weight matrix $W_i$ and a bias 
vector $b_i$: \[\mathbf{l_i} = f(W_i \cdot \mathbf{l_{i-1}} + b_i),\] 
where $f(\cdot)$ is a non-linearity like the sigmoid $f(x) = \frac{1}{1 
+ e^{-z}}$ applied component-wise.

An auto-encoder is such a neural network trained so as to minimize the 
error (specifically, the sum-squared difference) between the output 
layer and the input layer, over some set of input examples. If we have 
fewer nodes in the hidden layers than in the input or output layer, the 
neural network is forced to learn a compressed representation of the 
data (represented by the activations, or outputs, of the smaller number 
of hidden nodes). Instead, we keep a larger number of hidden nodes while 
imposing a sparsity constraint that the average activation of each one 
be small (e.g., 0.05). Other work has shown that such a constraint 
better forces the auto-encoder to learn interesting structures from the 
data~\cite{bengio}.

A normal auto-encoder only has one layer of hidden nodes, but we can 
also add more hidden layers to form a stacked auto-encoder. Per Bengio 
et al.~\shortcite{Bengio07greedylayer-wise}, we use a greedy strategy to 
train the extra hidden layers in the auto-encoder. We train an 
auto-encoder with one hidden layer as usual; once this completes, we 
obtain the output of the hidden layer over all the examples in the input 
set, and use that as the new set of examples. This allows us to learn 
higher-level representations of the data~\cite{Vincent}.

Other strategies for unsupervised feature learning include restricted 
Boltzmann machines (which are undirected), deep belief nets (stacked 
RBMs, akin to stacked auto-encoders), and denoising auto-encoders (which 
try to reconstruct some input from a corrupted version of it). We only 
evaluated sparse stacked auto-encoders in this paper.

The auto-encoder requires training data in the form of vectors where all 
the elements are between 0 and 1, or the range of the sigmoid function. 
We generated the training vectors by, for some chess board and for every 
type of piece (knight, rook, etc.), whether a white one (valued 1.0), a 
black one (valued 0.0), or none at all (valued 0.5) was present in every 
square of the chessboard.

\subsection{Manually-generated features}
\begin{table*}
\centering
\begin{tabular}{lp{0.6\textwidth}}
\hline
\textbf{Category} & \textbf{Description} \\ \hline
Position & Piece location, potential moves per piece, material count, 
square control. \\
Previous move & Capturing move, recapturing move, \emph{en passant} 
move, long castling move, short castling move, check move, stalemate 
move, checkmate move, promotion move. \\
\hline
\end{tabular} 
\caption{List of chess-specific manual features per position.}
\label{table:features}
\end{table*}

As a point of comparison to auto-encoders, we extracted chess-specific 
features manually from a given chess position (configuration of pieces 
on the board) and the move that resulted in that position; 
Table~\ref{table:features} contains a list. The position-based features 
include information such as where the pieces are on the board and 
potential moves, while the move-based features involve binary 
information such as if the move that resulted in this position was a 
capturing move or resulted in a check.  The resulting feature vector was 
then L2-normalized to fit the specifications of the support vector 
machine.

\section{Experimental methodology}
We processed all of the labeled training data to obtain a mapping from 
each unique game prefix (for most of the experiments, just the initial 
specified configuration of the board in the labeled game) to the labels 
that apply to it; most of the prefixes only had one label, while others 
had two or three. We then segmented the data (consisting of about 3000 
board configurations) into three disjoint sets: 90\% for training, 5\% 
as the validation set (for use with tuning our classifiers), and 5\% as 
the test set. The segmenting was done by random selection from the set 
of all labeled games.

In our experiments, we first use the labeled data in the training set 
(and possibly also unlabeled data) to train the classifiers. Then, for 
every test input chess game (or a prefix of a game), we use our 
classifiers to apply some number of labels to it that come from 
Table~\ref{table:themes}. We measure the performance of the predictions 
made by our classifiers by counting the number of true positives (a 
predicted label was among the set of gold labels for a test input), true 
negatives (a label in Table~\ref{table:themes} was neither in the set of 
predicted labels nor the set of gold labels), false positives (a 
predicted label was not in the set of gold labels), and false negatives 
(a gold label was not predicted). We compute and report the precision, 
recall, and F1 score using these measures of accuracy.

\subsection{Support vector machine classifier}
We note that any number of labels may apply to some test input that we 
want to classify; a straightforward use of a binary (or even multiclass) 
classifier, as provided by most implementations of support vector 
machines, will not suffice as it can only assign one label. Therefore, 
we take the one-versus-all strategy of training many binary SVMs; for 
each label that we consider, we train a binary classifier with all 
training games that have the given label as positive examples, and all 
other games as negative examples. We used {\tt 
SVMlight}~\cite{svmlight}, a popular third-party implementation of SVMs, 
for our experiments, using the manually-generated features extracted 
from each labeled data point.

\subsection{Classifying with the auto-encoder}
\label{sec:classifying-auto-encoder}
There exists several methods for using the auto-encoder for the task at 
hand. First, we can consider using the output of the last hidden layer 
as the feature vector for input into the SVM classifier, just as we did 
with the manually-generated features. To do this, we first perform 
unsupervised fine-tuning; we combine all the greedily-trained 
auto-encoders to make a many-layered neural network intended to 
reconstruct the input layer in its output layer, after which we again 
minimize the sum-squared error between the input and the output layers. 
We can then use the output of the hidden layer (i.e., not the one that 
reconstructs the input, but the middle layer which provides the most 
high-level representation of the input) as the feature vector.

We can also use semi-supervised learning methods, in which after 
greedily training the auto-encoders, we combine all the hidden layers of 
the auto-encoders and add an output layer at the end. For every labeled 
input game, we construct input-output vectors such that the output 
encodes the gold labels of the input. We then train the resulting 
multi-layer neural network to minimize the sum-squared error; this acts 
similarly to a multiclass logistic regression, except with a different 
objective at the last layer.

To train these neural networks, we used batch learning over all the 
examples at once. The loss function contained components for the 
sum-squared error as well as an L2 regularization term and a sparsity 
penalty based on the Kullback-Leibler divergence between the desired 
average activation of the hidden nodes and the actual average 
activation. We used backpropagation to compute the gradient of the loss 
function and L-BFGS to optimize the parameters as to minimize the loss 
function~\cite{ng-lecturenotes}. Running serially, the training takes a 
long time to complete, even on simple inputs. Therefore, we examined 
ways to parallelize the task in order to fully leverage the capacity of 
modern many-core machines; we achieved a significant speedup by 
parallelizing the loss function evaluation to 16 execution threads.

\subsection{Random baseline}
To use as a baseline and ensure that our classifiers are giving sensible 
results, we randomly made predictions for every test game, where each 
one assigned each possible label at random, with the possibility of the 
label being given based on the prior probability of the label, or the 
proportion of the training games which have that label. The random 
predictions were made 1,000 times, with the predictions made 
independently at random, and the error rates were averaged.

\section{Experimental results}
\subsection{SVM classifier}
\begin{table*}
\centering
\begin{tabular}{lccccccc}
\hline
\textbf{Type} & \textbf{TP} & \textbf{FP} & \textbf{TN} & \textbf{FN} & 
\textbf{Precision} & \textbf{Recall} & \textbf{F1} \\ \hline
Unsegmented & 0.21\% & 0.00\% & 96.89\% & 2.90\% & 100\% & 6.71\% & 
12.57\% \\
Baseline (Random) & 0.42\% & 2.76\% & 94.12\% & 2.69\% & 13.32\% & 
13.65\% & 13.48\% \\
\hline
Segmented & 0.13\% & 0.02\% & 98.34\% & 1.51\% & 86.67\% & 7.65\% & 
14.05\% \\
Baseline (Random) & 0.12\% & 1.49\% & 96.87\% & 1.51\% & 7.65\% & 7.55\% 
& 7.60\% \\ \hline
\end{tabular} 
\caption{Classification results for the games that have not been 
segmented by the color of the active player, and those that have been. 
In the former, precision clearly outperforms a random baseline, but 
recall falls short, resulting in a lower F1 score. In the latter, 
precision has fallen relative to the unsegmented version, but recall has 
risen, resulting in a slightly higher F1 score. In contrast, the random 
model became much worse in both ways.}
\label{table:svmResults}
\end{table*}

We first describe the results achieved when we trained using the 
manually-specified feature extractors. Since each game has very few 
labels, most of the binary classification results are negative. 
Therefore, the model is heavily biased toward negative results, as shown 
in Table~\ref{table:svmResults}.

This classification method had relatively high precision compared to the 
random baseline. However, the recall is much lower than that of the 
random classifier. As a result, the F1 score (a combination of the 
precision and recall scores) of the SVM-based model is no better than 
that of the random model.

When we extended the labeling to also incorporate the first mover's 
color information, the SVM model's recall was improved at a minor cost 
to precision, resulting in a higher F1 score. In contrast, the random 
predictions became much worse: Both the precision and recall scores fell 
as a result of this change, which reduced the evenness of the label 
distribution among the training games, and reduced the amount of games 
with each label (since games are now spread over twice as many tags).
\begin{table}
\centering
\begin{tabular}{lr}
\hline
\textbf{Tag} & \textbf{Error rate} \\ \hline
\tt{queenCapture}  & 0.6452\% \\
\tt{kingInCenter}  & 0.6452\% \\
\tt{perpetualCheck}  & 0.6452\% \\
\tt{drawCombinationsStalemate}  & 3.8710\% \\
\tt{bishopPair}  & 0.6452\% \\
\tt{doubleAttack}  & 0.6452\% \\
\tt{clearance}  & 2.5806\% \\
\tt{openingClosingDiagonals}  & 0.6452\% \\
\tt{openingClosingFiles}  & 0.6452\% \\
\tt{enclosedKings}  & 1.2903\% \\
\tt{discoveredAttack}  & 2.5806\% \\
\tt{passedPawn}  & 1.2903\% \\
\tt{quietMates}  & 3.8710\% \\
\tt{pin}  & 5.1613\% \\
\tt{removalOfDefence}  & 22.5806\% \\
\tt{deflection}  & 25.1613\% \\
\tt{decoy}  & 9.6774\% \\
\tt{interference}  & 3.8710\% \\
\tt{backrankWeakness}  & 3.2258\% \\
\tt{freeingBlockingSquares}  & 1.9355\% \\
\tt{pawnPromotion}  & 6.4516\% \\
\tt{pullingPiecesToBadSquares}  & 0.6452\% \\ \hline
\end{tabular} 
\caption{Percent of incorrect classification (false negatives and false 
positives) for various tags.}
\label{table:negatives}
\end{table}

Given the high precision but low recall, we delved deeper to analyze the 
sources of errors and true positives. As we show in 
Table~\ref{table:negatives}, we found that {\tt deflection} and {\tt 
removalOfDefence} are much more likely to cause false negatives than 
other tags. In addition, we found that all of the true positives are for 
the {\tt quietMates} tag. 9 of the 11 true positive {\tt quietMates} 
games were classified correctly using only the features that describe 
the contents of the board. On the other hand, 5 true positive {\tt 
quietMates} games were correctly classified when using only features 
about the possible moves in that position. Interestingly, disabling the 
``checkmate move'' feature did not affect the number of true positives.

\subsection{Classifying with the auto-encoder}
\begin{table*}
\centering
\begin{tabular}{ccccccccc}
\hline
\textbf{Type} & \textbf{Hidden layers} & \textbf{TP} & \textbf{FP} & 
\textbf{TN} & \textbf{FN} & \textbf{Precision} & \textbf{Recall} & 
\textbf{F1} \\ \hline
Unsegmented & 0 & 1.49\% & 3.62\% & 93.26\% & 1.61\% & 29.26\% & 48.17\% 
& 36.40\% \\
Unsegmented & 1 & 1.35\% & 3.83\% & 93.06\% & 1.76\% & 26.01\% & 43.29\% 
& 32.49\% \\
Unsegmented & 2 & 1.40\% & 4.48\% & 92.41\% & 1.71\% & 23.87\% & 45.12\% 
& 31.22\% \\
Unsegmented & 3 & 1.40\% & 4.48\% & 92.41\% & 1.71\% & 23.87\% & 45.12\% 
& 31.22\% \\
\hline
Segmented & 0 & 0.38\% & 1.01\% & 97.35\% & 1.26\% & 27.08\% & 22.94\% & 
24.84\% \\
Segmented & 1 & 0.49\% & 2.90\% & 95.46\% & 1.15\% & 14.49\% & 30.00\% & 
19.54\% \\
Segmented & 2 & 0.22\% & 1.27\% & 97.09\% & 1.42\% & 14.84\% & 13.53\% & 
14.15\% \\
Segmented & 3 & 0.59\% & 3.89\% & 94.47\% & 1.05\% & 13.12\% & 35.88\% & 
19.21\% \\ \hline
\end{tabular} 
\caption{Auto-encoder classification results for the games that have not 
been segmented by the color of the active player, and those that have 
been, as the number of hidden layers varies.}
\label{table:auto-encoder-results}
\end{table*}

We first tried to use the unsupervised fine-tuned stacked auto-encoder's 
hidden outputs as the input to the SVM classifiers, as with the 
manually-generated feature vectors. However, we found that the resulting 
feature vectors were not very useful for SVM, as the classifiers did not 
predict any labels as true for all training or test input games.

Instead, we used the semi-supervised classification method as described 
in Section~\ref{sec:classifying-auto-encoder}. First, we greedily train 
each hidden layer as an auto-encoder on the training, test, and 
validation sets. We trained the combined neural network composed of 
greedily-trained auto-encoders to, for each input, predict an output 
vector corresponding to the labels that apply to the input---0 if a 
label did not apply, and 1 if it did. From the output vectors computed 
with the elements in the validation set, we tuned the threshold we use 
in considering whether a label is present or not present to maximize the 
F1 score. In Table~\ref{table:auto-encoder-results}, we report the 
performance achieved on the test set when using the threshold tuned to 
maximize the F1 score on the validation set.

These results show higher F1 score than the SVM classifier, with 
significantly lower precision but also much higher recall. 
Unfortunately, the results indicate that the representation of the board 
learned by the auto-encoder does not correlate very well with the 
concepts we wanted to identify, given that having no hidden layers 
resulted in higher scores. Adding extra hidden layers seems to not have 
helped improve the classification results, except in the last instance, 
where an extra layer improved the F1 score to original levels. It is 
hard to tell whether some slight change in the auto-encoder training 
procedure will be able to drastically improve the results, or whether 
significant changes in the architecture will be required. We also were 
not able to use the unlabeled chess games we collected with the 
auto-encoder, due to the long training time required and our relatively 
slow implementation of its training.

\subsection{Analysis}
Given the nature of labeled training data and their assembly from 
disparate sources, there exists some redundancy among the labels, and 
games may not have all the labels that they should have in our 
chess-description ontology. Furthermore, many of the labels had very few 
(fewer than twenty) games as positive examples, and performance may 
improve greatly if we had a larger training set to use.

\section{Future Work}
We identify three areas in which future work can make progress on 
automated sportscasting. Using a higher-performance auto-encoder 
training procedure and training it on a larger data set could help 
improve the results. We could also extend our system with an automated 
text-generation layer. In addition, leveraging chess-specific tools and 
features like chess engines and evaluation tools could help improve the 
quality of the output. Alternatively, an orthogonal path of applying 
similar general approaches to other games and attempting to learn 
higher-level constructs, including complex ones like StarCraft and 
simpler ones like Checkers, merits exploration.

\subsection{Performance}
One of the main limitations of our work is performance. Due to the slow 
performance of the auto-encoder, we were not able to fully exploit the 
millions of games that we have available to generate features for our 
unsupervised system. Therefore, extensions to our work should 
investigate ways to improve the performance so that the auto-encoder can 
process more inputs and thus generate more relevant feature functions.

\paragraph{Distributed training.}
Our auto-encoder takes advantage of some of the parallel processing 
capacity of the host system through a local MapReduce scheme. Extending 
that with, for example, a distributed MapReduce framework like Hadoop 
may improve performance if the additional cost of network I/O are 
outweighed by the benefits of parallelization.

\paragraph{Using the GPU.}
Another option that we partially explored is the exploitation of the 
processing power of the graphics processing unit (GPU). Function 
optimization, as used in training the auto-encoder, is a 
computation-heavy task, and the GPU is specifically designed for 
massively-parallel floating point computations. We investigated using 
JavaCL, an interface to using the GPU for general purpose computation, 
to parallelize the auto-encoder, and demonstrated that it is definitely 
superior at parallelizable and computation-intensive tasks like matrix 
multiplication, because the benefits of greater throughput through 
greater parallelization outweigh the costs of memory transfer.

\subsection{Text Generation and Linguistic Grounding}
One further avenue of exploration is a system that automatically 
generates text given a chess game with tagged positions. The naive 
approach would be to build a template-based text generation system, 
which could be as simple as generating a formatted sentence highlighting 
hand-selected salient features for each tag or something more 
complicated such as a context-free grammar. In the context of this 
project, we found that some chess positions were positive examples of 
multiple tags. As such, another possible avenue of research could be to 
conglomerate the annotations for each specific position and generate an 
annotation that makes sense in its context.

While we did use high-level structures as defined by experts and 
discussed in detailed commentary of chess games, we did not attempt to 
learn and mimic the actual structures used in the natural language 
commentary. To have a useful commentating system for an intricate game 
like chess, we need to both discover the high-level structures in the 
game, as well as know how to talk about the high-level structures. 
Extending the approach taken by some of the automated sportscasting work 
discussed in the Related Works section, it may be possible to learn both 
high-level concepts that are present in a game, and how to realize 
descriptions of them in language.

\subsection{Improving Results}
In pursuing this project, we attempted to maintain an approach that is 
as general as possible so that it can be applied to other contexts, and 
to minimize the amount of engineering required for components that are 
very specific to chess. Given the complexity of chess, a relatively 
straightforward, but chess-specific, way to improve the performance 
would be to incorporate tools like chess engines. Computer chess has 
been an active field of research, and much work has gone into game 
engines and weight functions which excel at the game.

For example, a chess-specific rule-based approach should be able to 
easily identify the more precisely-described themes like {\tt 
doubleAttack} and {\tt queenCapture} by analyzing the future branches of 
the game and looking for specific outcomes. However, this approach would 
require that each theme must have a specific heuristic function written 
for that express purpose to identify certain properties in future states 
in order to effectively identify tags using a chess engine. In addition, 
it may be difficult to identify more subjective themes like {\tt 
clearance} and {\tt kingToBadSquares} that are based more on human 
judgement.

This suggests a combined approach which can leverage the strengths of 
each approach by using the chess-specific features when they are 
available, and combining that with results from statistical learning. 
For instance, while the chess engine can easily identify imminent mate 
combinations, it would require a coordination of both the engine and the 
statistical learning tool to identify threshold functions like {\tt 
kingToBadSquares}. Machine learning would be used to optimize weights 
and thresholds (e.g. for whether a king's position is bad) while the 
engine would be used to explore future paths. For example, a deep 
traversal of many future paths by a chess engine may reveal that an 
attack is possible ({\tt attackPossible}) or that the player can force a 
pawn promotion ({\tt pawnPromotion}).

\subsection{Generalization to Other Games}
One goal of this project was to develop a statistical machine learning 
approach that was general enough to be easily adapted to other 
applications like StarCraft or Checkers. Given the light dependence on 
chess-specific concepts of the approach described in this paper, 
specifically with the auto-encoder, we should be able to easily adapt 
this to other turn-based games.

\section{Conclusion}
We built two statistical learning systems to classify high-level 
positional concepts for chess, and described various issues involved in 
implementing these systems. Both classifiers are feature driven and 
classified by a support vector machine. The first system uses manually 
generated features for a chessboard configuration and the move that led 
to it, while the second system uses a sparse auto-encoder to 
automatically discover high-level structures (and hence features) for a 
given chessboard configuration in an unsupervised fashion. We found that 
the auto-encoder-based classifier outperformed the manual feature 
vector-based classifier, which outperformed the random baseline 
classifier: on the same segmented data set, the random baseline 
classifier had an F1 score of $\mathbf{7.60}$, the best manual feature 
vector-based classifier had an F1 score of $\mathbf{14.05}$, and the 
best auto-encoder-based classifier had an F1 score of $\mathbf{24.84}$.

\section{Acknowledgements}
We would like to thank Dan Klein for teaching us statistical machine 
learning techniques, Darren Kuo for helping with initial planning and 
game de-duplication, and Soda Hall for accommodating us for many days 
and nights.

\bibliographystyle{acl}
\bibliography{paper}

\end{document}